Ethereum: A Secure Decentralised Generalised Transaction Ledger \\ {\smaller \textbf{EIP-150 revision (\YellowPaperVersionNumber{})}}

### 2017-09-29 Paper.tex
Cross references, grammar, maths in Appendix I and J, and section 11

\documentclass[9pt,oneside]{amsart}
\usepackage{cancel}
%\usepackage{tweaklist}
\usepackage{xspace}
\usepackage{graphicx}
\usepackage{multicol}
\usepackage{subfig}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage[a4paper,width=170mm,top=18mm,bottom=22mm,includeheadfoot]{geometry}
\usepackage{booktabs}
\usepackage{array}
\usepackage{verbatim}
\usepackage{caption}
\usepackage{natbib}
\usepackage{float}
\usepackage{pdflscape}
\usepackage{mathtools}
\usepackage[usenames,dvipsnames]{xcolor}
\usepackage{afterpage}
\usepackage{tikz}
\usepackage{underscore}
%\usepackage{lipsum} % to generate filler text for testing of document rendering
\usepackage[bookmarks=true, unicode=true, pdftitle={Ethereum Yellow Paper: a formal specification of Ethereum, a programmable blockchain}, pdfauthor={Gavin Wood and others as per https://github.com/ethereum/yellowpaper/commits/master},pdfkeywords={Ethereum, Yellow Paper, blockchain, virtual machine, cryptography, decentralised, singleton, transaction, generalised},pdfborder={0 0 0.5 [1 3]}]{hyperref}
%,pagebackref=true

@@

@@

\hypertarget{Pi}{}
Formally, we expand to:
\begin{eqnarray}
\boldsymbol{\sigma}_{t+1} & \equiv & \Pi(\boldsymbol{\sigma}_\mathbf{t}, B) \\
B & \equiv & (..., (T_0, T_1, ...) ) \\
\Pi(\boldsymbol{\sigma}, B) & \equiv & \hyperlink{Omega}{\Omega}(B, \hyperlink{Upsilon}{\Upsilon}(\Upsilon(\boldsymbol{\sigma}, T_0), T_1) ...)
\end{eqnarray}

Where \hyperlink{Omega}{$\Omega$} is the block-finalisation state transition function (a function that rewards a nominated party); $B$ is this block, which includes a series of transactions amongst some other components; and $\Pi$ is the block-level state-transition function.

@@

Since the system is decentralised and all parties have an opportunity to create a new block on some older pre-existing block, the resultant structure is necessarily a tree of blocks. In order to form a consensus as to which path, from root (the genesis block) to leaf (the block containing the most recent transactions) through this tree structure, known as the blockchain, there must be an agreed-upon scheme. If there is ever a disagreement between nodes as to which root-to-leaf path down the block tree is the `best' blockchain, then a \textit{fork} occurs.

@@

\begin{description}\hypertarget{H p}{
\item[parentHash] The Keccak 256-bit hash of the parent block's header, in its entirety; formally $H_\mathbf{p}$.}
\item[ommersHash] The Keccak 256-bit hash of the ommers list portion of this block; formally $\hyperlink{H o}{H_\mathbf{o}}$.\hypertarget{H c}{
\item[beneficiary] The 160-bit address to which all fees collected from the successful mining of this block shall be transferred; formally $H_\mathbf{c}$.}
\item[stateRoot] The Keccak 256-bit hash of the root node of the state \hyperlink{trie}{trie}, after all transactions are executed and finalisations applied; formally \hyperlink{H r}{$H_\mathbf{r}$}.
\item[transactionsRoot] The Keccak 256-bit hash of the root node of the \hyperlink{trie}{trie} structure populated with each transaction in the transactions list portion of the block; formally \hyperlink{H t}{$H_\mathbf{t}$}.
\item[receiptsRoot] The Keccak 256-bit hash of the root node of the \hyperlink{trie}{trie} structure populated with the receipts of each transaction in the transactions list portion of the block; formally \hyperlink{H e}{$H_\mathbf{e}$}.
\item[logsBloom] The Bloom filter composed from indexable information (logger address and log topics) contained in each log entry from the receipt of each transaction in the transactions list; formally \hyperlink{H b}{$H_\mathbf{b}.$}\hypertarget{H d}{
\item[difficulty] A scalar value corresponding to the difficulty level of this block. This can be calculated from the previous block's difficulty level and the timestamp; formally $H_\mathbf{d}$}.\hypertarget{H i}{
\item[number] A scalar value equal to the number of ancestor blocks. The genesis block has a number of zero; formally $H_\mathbf{i}$.}\hypertarget{H l}{
\item[gasLimit] A scalar value equal to the current limit of gas expenditure per block; formally $H_\mathbf{l}$.}\hypertarget{H g}{
\item[gasUsed] A scalar value equal to the total gas used in transactions in this block; formally $H_\mathbf{g}$.}
\item[timestamp] A scalar value equal to the reasonable output of Unix's time() at this block's inception; formally \hyperlink{H s}{$H_\mathbf{s}$}.\hypertarget{H x}{
\item[extraData] An arbitrary byte array containing data relevant to this block. This must be 32 bytes or fewer; formally $H_\mathbf{x}$.}\hypertarget{H m}{
\item[mixHash] A 256-bit hash which proves combined with the nonce that a sufficient amount of computation has been carried out on this block; formally $H_\mathbf{m}$.}
\hypertarget{block nonce}{\hypertarget{H n}{
\item[nonce] A 64-bit hash which proves combined with the mix-hash that a sufficient amount of computation has been carried out on this block; formally $H_\mathbf{n}$.}}
\end{description}
\hypertarget{B U}{
The other two components in the block are simply a list of ommer block headers (of the same format as above), $B_\mathbf{U}$, and a series of the transactions, $B_\mathbf{T}$. Formally, we can refer to a block $B$:
\begin{equation}
B \equiv (B_H, B_\mathbf{T}, B_\mathbf{U})
\end{equation}}

@@

The transaction receipt is a tuple of four items \hypertarget{R sigma}{comprising the post-transaction state, $R_{\boldsymbol{\sigma}}$}, \hypertarget{R u}{the cumulative gas used in the block containing the transaction receipt as of immediately after the transaction has happened, $R_\mathbf{u}$}, the set of logs created through execution of the transaction, \hyperlink{R l}{$R_\mathbf{l}$} and the Bloom filter composed from information in those logs, \hyperlink{R b}{$R_\mathbf{b}$}:
\begin{equation}
R \equiv (R_{\boldsymbol{\sigma}}, R_\mathbf{u}, \hyperlink{R b}{R_\mathbf{b}}, \hyperlink{R l}{R_\mathbf{l}})
\end{equation}

@@

We define the Bloom filter function, $M$, to reduce a log entry into a single 256-byte hash:
\begin{equation}
M(O) \equiv \bigvee_{t \in \{O_\mathbf{a}\} \cup O_\mathbf{t}} \big( M_{3:2048}(t) \big)
\end{equation}

where $M_{3:2048}$ is a specialised Bloom filter that sets three bits out of 2048, given an arbitrary byte sequence. It does this through taking the low-order 11 bits of each of the first three pairs of bytes in a Keccak-256 hash of the byte sequence./footnote{11 bits $= 2^2048$, and the low-order 11 bits is the modulo 2048 of the operand, which is in this case is "each of the first three pairs of bytes in a Keccak-256 hash of the byte sequence."} Formally:
\begin{eqnarray}
M_{3:2048}(\mathbf{x}: \mathbf{x} \in \mathbb{B}) & \equiv & \mathbf{y}: \mathbf{y} \in \mathbb{B}_{256} \quad \text{where:}\\
\mathbf{y} & = & (0, 0, ..., 0) \quad \text{except:}\\
\forall_{i \in \{0, 2, 4\}}&:& \mathcal{B}_{m(\mathbf{x}, i)}(\mathbf{y}) = 1\\
m(\mathbf{x}, i) &\equiv& \mathtt{\tiny KEC}(\mathbf{x})[i, i + 1] \bmod 2048
\end{eqnarray}

@@

We can assert a block's validity if and only if it satisfies several conditions: it must be internally consistent with the ommer and transaction block hashes and the given transactions $B_\mathbf{T}$ (as specified in sec \ref{ch:finalisation}), when executed in order on the base state $\boldsymbol{\sigma}$ (derived from the final state of the parent block), result in a new state of the identity $H_\mathbf{r}$:
\begin{equation}
\begin{array}[t]{lclc}\hypertarget{H r}{
H_\mathbf{r} &\equiv& \mathtt{\small \hyperlink{trie}{TRIE}}(L_S(\hyperlink{Pi}{\Pi}(\boldsymbol{\sigma}, B))) & \wedge }\hypertarget{H o}{\\
H_\mathbf{o} &\equiv& \mathtt{\small KEC}(\mathtt{\small RLP}(L_H^*(\hyperlink{B U}{B_\mathbf{U}})))} & \wedge \hypertarget{H t}{\\
H_\mathbf{t} &\equiv& \mathtt{\small TRIE}(\{\forall i < \lVert B_\mathbf{T} \rVert, i \in \mathbb{P}: p(i, L_\mathbf{T}(B_\mathbf{T}[i]))\}) }& \wedge \hypertarget{H e}{\\
H_\mathbf{e} &\equiv& \mathtt{\small TRIE}(\{\forall i < \lVert B_\mathbf{R} \rVert, i \in \mathbb{P}: p(i, L_R(B_\mathbf{R}[i]))\})} & \wedge \hypertarget{H b}{\\
H_\mathbf{b} &\equiv& \bigvee_{\mathbf{r} \in B_\mathbf{R}} \big( \mathbf{r}_\mathbf{b} \big)}
\end{array}
\end{equation}

@@

The values stemming from the computation of transactions, specifically the transaction receipts, $B_\mathbf{R}$, and that defined through the transaction's \hyperlink{Pi}{state-accumulation function, $\Pi$,} are formalised later in section \ref{sec:statenoncevalidation}.

@@

The nonce of a block, $\hyperlink{H n}{H_\mathbf{n}}$, must satisfy the relations:
\begin{equation}
n \leqslant \frac{2^{256}}{\hyperlink{H d}{H_\mathbf{d}}} \quad \wedge \quad \mathbf{m} = \hyperlink{H m}{H_\mathbf{m}}
\end{equation}
with $(n, m) = \mathtt{PoW}(H_{\hcancel{n}}, \hyperlink{H n}{H_\mathbf{n}}, \mathbf{d})$.

@@

Thus we are able to define the \hypertarget{BlockHeaderValidityFunction}{block header validity function $V(H)$}:\label{eq:BlockHeaderValidityFunction}
\begin{eqnarray}
V(H) & \equiv &  n \leqslant \frac{2^{256}}{\hyperlink{H d}{H_\mathbf{d}}} \wedge m = \hyperlink{H m}{H_\mathbf{m}} \quad \wedge \\
& & \hyperlink{H d}{H_\mathbf{d}} = D(H) \quad \wedge \\
& & \hyperlink{H g}{H_\mathbf{g}} \le \hyperlink{H l}{H_\mathbf{l}}  \quad \wedge \\
& & \hyperlink{H l}{H_\mathbf{l}} < {P(H)_H}_\mathbf{l} + \left\lfloor\frac{{P(H)_H}_\mathbf{l}}{1024}\right\rfloor  \quad \wedge \\
& & \hyperlink{H l}{H_\mathbf{l}} > {P(H)_H}_\mathbf{l} - \left\lfloor\frac{{P(H)_H}_\mathbf{l}}{1024}\right\rfloor  \quad \wedge \\
& & \hyperlink{H l}{H_\mathbf{l}} \geqslant 125000  \quad \wedge \\
& & \hyperlink{H s}{H_\mathbf{s}} > {P(H)_H}_\mathbf{s} \quad \wedge \\
& & \hyperlink{H i}{H_\mathbf{i}} = {P(H)_H}_\mathbf{i} +1 \quad \wedge \\
& & \lVert \hyperlink{H x}{H_\mathbf{x}} \rVert \le 32
\end{eqnarray}
where $(\mathbf{n}, \mathbf{m}) = \mathtt{PoW}(H_{\hyperlink{h cancel n}{\hcancel{n}}}, \hyperlink{H n}{H_\mathbf{n}}, \mathbf{d})$

@@

The canonical blockchain is a path from root to leaf through the entire block tree. In order to have consensus over which path it is, conceptually we identify the path that has had the most computation done upon it, or, the \textit{heaviest} path. Clearly one factor that helps determine the heaviest path is the block number of the leaf, equivalent to the number of blocks, not counting the unmined genesis block, in the path. The longer the path, the greater the total mining effort that must have been done in order to arrive at the leaf. This is akin to existing schemes, such as that employed in Bitcoin-derived protocols.

@@

\hypertarget{Gamma}{
We may now define the function, $\Gamma$, that maps a block $B$ to its initiation state:
\begin{equation}
\Gamma(B) \equiv \begin{cases}
\boldsymbol{\sigma}_0 & \text{if} \quad P(B_H) = \varnothing \\
\boldsymbol{\sigma}_\mathbf{i}: \mathtt{\small \hyperlink{trie}{TRIE}}(L_S(\boldsymbol{\sigma}_\mathbf{i})) = {P(B_H)_H}_\mathbf{r} & \text{otherwise}
\end{cases}
\end{equation}
}
Here, $\mathtt{\small TRIE}(L_S(\boldsymbol{\sigma}_\mathbf{i}))$ means the hash of the root node of a trie of state $\boldsymbol{\sigma}_\mathbf{i}$; it is assumed that implementations will store this in the state database, which is trivial and efficient since the trie is by nature an immutable data structure.
\hypertarget{Phi}{
And finally we define $\Phi$, the block transition function, which maps an incomplete block $B$ to a complete block $B'$:}
\begin{eqnarray}
\Phi(B) & \equiv & B': \quad B' = B^* \quad \text{except:} \\
B'_\mathbf{n} & = & n: \quad x \leqslant \frac{2^{256}}{\hyperlink{H d}{H_\mathbf{d}}} \\
B'_\mathbf{m} & = & m \quad \text{with } (x, m) = \mathtt{PoW}(B^*_{\hyperlink{h cancel n}{\hcancel{n}}}, n, \mathbf{d}) \\
B^* & \equiv & B \quad \text{except:} \quad B^*_\mathbf{r} = r(\hyperlink{Pi}{\Pi}(\Gamma(B), B))
\end{eqnarray}
With $\mathbf{d}$ being a dataset as specified in appendix \ref{app:ethash}.

@@

In the case of $B_\mathbf{R}[n]_\mathbf{u}$, we take a similar approach defining each item as the gas used in evaluating the corresponding transaction summed with the previous item (or zero, if it is the first), giving us a running total:
\begin{equation}
\mathbf{R}[n]_\mathbf{u} = \begin{cases} 0 & \text{if} \quad n < 0 \\
\begin{array}[b]{l}
\hyperlink{Upsilong}{\Upsilon^g}(\mathbf{R}[n - 1]_{\boldsymbol{\sigma}}, B_\mathbf{T}[n])\\ \quad + \mathbf{R}[n-1]_\mathbf{u}
\end{array}
 & \text{otherwise} \end{cases}
\end{equation}

@@

Finally, we define \hyperlink{Pi}{$\Pi$} as the new state given the block reward function \hyperlink{Omega}{$\Omega$} applied to the final transaction's resultant state, $\hyperlink{ell}{\ell}(B_\mathbf{R})_{\boldsymbol{\sigma}}$:
\begin{equation}
\Pi(\boldsymbol{\sigma}, B) \equiv \hyperlink{Omega}{\Omega}(B, \ell(\hyperlink{R sigma}{\mathbf{R})_{\boldsymbol{\sigma}}})
\end{equation}

@@

One plague of the Bitcoin world is ASICs. These are specialised pieces of compute hardware that exist only to do a single task. In Bitcoin's case the task is the SHA256 hash function. While ASICs exist for a proof-of-work function, both goals are placed in jeopardy. Because of this, a proof-of-work function that is ASIC-resistant (i.e. difficult or economically inefficient to implement in specialised compute hardware) has been identified as the proverbial silver bullet.

Two directions exist for ASIC resistance; firstly make it sequential memory-hard, i.e. engineer the function such that the determination of the \hyperlink{block nonce}{nonce} requires a lot of memory and bandwidth such that the memory cannot be used in parallel to discover multiple nonces simultaneously. The second is to make the type of computation it would need to do general-purpose; the meaning of ``specialised hardware''  for a general-purpose task set is, naturally, general purpose hardware and as such commodity desktop computers are likely to be pretty close to ``specialised hardware'' for the task. For Ethereum 1.0 we have chosen the first path.

More formally, the proof-of-work function takes the form of $\mathtt{PoW}$:
\begin{equation}
m = H_\mathbf{m} \quad \wedge \quad \mathbf{n} \leqslant \frac{2^{256}}{\hyperlink{H d}{H_\mathbf{d}}} \quad \text{with} \quad (\mathbf{m}, \mathbf{n}) = \mathtt{PoW}(H_{\hyperlink{h cancel n}{\hcancel{n}}}, H_\mathbf{n}, \mathbf{d})
\end{equation}

Where $H_{\hyperlink{h cancel n}{\hcancel{n}}}$ is the new block's header but \textit{without} the nonce and mix-hash components; $\hyperlink{H n}{H_\mathbf{n}}$ is the nonce of the block header; $\mathbf{d}$ is a large data set needed to compute the mixHash and $\hyperlink{H d}{H_\mathbf{d}}$ is the new block's difficulty value (i.e. the block difficulty from section \ref{ch:ghost}). $\mathtt{PoW}$ is the proof-of-work function which evaluates to an array with the first item being the mixHash and the second item being a pseudo-random number cryptographically dependent on $H$ and $\mathbf{d}$. The underlying algorithm is called Ethash and is described below.
\subsubsection{Ethash}
Ethash is the PoW algorithm for Ethereum 1.0. It is the latest version of Dagger-Hashimoto, introduced by \cite{dagger} and \cite{hashimoto}, although it can no longer appropriately be called that since many of the original features of both algorithms have been drastically changed in the last month of research and development. The general route that the algorithm takes is as follows:

>>>>

\documentclass[9pt,oneside]{amsart}
\usepackage{cancel}
%\usepackage{tweaklist}
\usepackage{xspace}
\usepackage{graphicx}
\usepackage{multicol}
\usepackage{subfig}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage[a4paper,width=170mm,top=18mm,bottom=22mm,includeheadfoot]{geometry}
\usepackage{booktabs}
\usepackage{array}
\usepackage{verbatim}
\usepackage{caption}
\usepackage{natbib}
\usepackage{float}
\usepackage{pdflscape}
\usepackage{mathtools}
\usepackage[usenames,dvipsnames]{xcolor}
\usepackage{afterpage}
\usepackage{tikz}
\usepackage{underscore}
%\usepackage{lipsum} % to generate filler text for testing of document rendering
\usepackage[english]{babel}
\usepackage [autostyle]{csquotes}
\MakeOuterQuote{"}
\usepackage[bookmarks=true, unicode=true, pdftitle={Ethereum Yellow Paper: a formal specification of Ethereum, a programmable blockchain}, pdfauthor={Gavin Wood and others as per https://github.com/ethereum/yellowpaper/commits/master},pdfkeywords={Ethereum, Yellow Paper, blockchain, virtual machine, cryptography, decentralised, singleton, transaction, generalised},pdfborder={0 0 0.5 [1 3]}]{hyperref}
%,pagebackref=true

@@

%\usepackage{glossaries}

@@

Formally, we expand to:
\begin{eqnarray}
\boldsymbol{\sigma}_{t+1} & \equiv & \hyperlink{Pi}{\Pi}(\boldsymbol{\sigma}_\mathbf{t}, B) \\
B & \equiv & (..., (T_0, T_1, ...) ) \\
\Pi(\boldsymbol{\sigma}, B) & \equiv & \hyperlink{Omega}{\Omega}(B, \hyperlink{Upsilon}{\Upsilon}(\Upsilon(\boldsymbol{\sigma}, T_0), T_1) ...)
\end{eqnarray}

Where \hyperlink{Omega}{$\Omega$} is the block-finalisation state transition function (a function that rewards a nominated party); $B$ is this block, which includes a series of transactions amongst some other components; and $\hyperlink{Pi}{\Pi}$ is the block-level state-transition function.

@@

Since the system is decentralised and all parties have an opportunity to create a new block on some older pre-existing block, the resultant structure is necessarily a tree of blocks. In order to form a consensus as to which path, from root (the \hyperlink{GenesisBlock}{genesis block}) to leaf (the block containing the most recent transactions) through this tree structure, known as the blockchain, there must be an agreed-upon scheme. If there is ever a disagreement between nodes as to which root-to-leaf path down the block tree is the `best' blockchain, then a \textit{fork} occurs.

@@

\begin{description}\hypertarget{H p}{
\item[parentHash] The Keccak 256-bit hash of the parent block's header, in its entirety; formally $H_\mathbf{p}$.}
\item[ommersHash] The Keccak 256-bit hash of the ommers list portion of this block; formally $\hyperlink{H o}{H_\mathbf{o}}$.\hypertarget{H c}{
\item[beneficiary] The 160-bit address to which all fees collected from the successful mining of this block shall be transferred; formally $H_\mathbf{c}$.}
\item[stateRoot] The Keccak 256-bit hash of the root node of the state \hyperlink{trie}{trie}, after all transactions are executed and finalisations applied; formally \hyperlink{H r}{$H_\mathbf{r}$}.
\item[transactionsRoot] The Keccak 256-bit hash of the root node of the \hyperlink{trie}{trie} structure populated with each transaction in the transactions list portion of the block; formally \hyperlink{H t}{$H_\mathbf{t}$}.
\item[receiptsRoot] The Keccak 256-bit hash of the root node of the \hyperlink{trie}{trie} structure populated with the receipts of each transaction in the transactions list portion of the block; formally \hyperlink{H e}{$H_\mathbf{e}$}.
\item[logsBloom] The Bloom filter composed from indexable information (logger address and log topics) contained in each log entry from the receipt of each transaction in the transactions list; formally \hyperlink{H b}{$H_\mathbf{b}.$}\hypertarget{H d}{
\item[difficulty] A scalar value corresponding to the difficulty level of this block. This can be calculated from the previous block's difficulty level and the timestamp; formally $H_\mathbf{d}$}.\hypertarget{H i}{
\item[number] A scalar value equal to the number of ancestor blocks. The \hyperlink{GenesisBlock}{genesis block} has a number of zero; formally $H_\mathbf{i}$.}\hypertarget{H l}{
\item[gasLimit] A scalar value equal to the current limit of gas expenditure per block; formally $H_\mathbf{l}$.}\hypertarget{H g}{
\item[gasUsed] A scalar value equal to the total gas used in transactions in this block; formally $H_\mathbf{g}$.}\hypertarget{H s}{
\item[timestamp] A scalar value equal to the reasonable output of Unix's time() at this block's inception; formally {$H_\mathbf{s}$}.}\hypertarget{H x}{
\item[extraData] An arbitrary byte array containing data relevant to this block. This must be 32 bytes or fewer; formally $H_\mathbf{x}$.}\hypertarget{H m}{
\item[mixHash] A 256-bit hash which proves combined with the nonce that a sufficient amount of computation has been carried out on this block; formally $H_\mathbf{m}$.}
\hypertarget{block nonce}{\hypertarget{H n}{
\item[nonce] A 64-bit hash which proves combined with the mix-hash that a sufficient amount of computation has been carried out on this block; formally $H_\mathbf{n}$.}}
\end{description}
\hypertarget{B U}{
The other two components in the block are simply a list of ommer block headers (of the same format as above), $B_\mathbf{U}$, and a series of the transactions, $B_\mathbf{T}$. Formally, we can refer to a block $B$:
\begin{equation}
B \equiv (B_H, B_\mathbf{T}, B_\mathbf{U})
\end{equation}}

@@

\hypertarget{R}{}
The transaction receipt, $R$, is a tuple of four items \hypertarget{R sigma}{comprising the post-transaction state, $R_{\boldsymbol{\sigma}}$}, \hypertarget{R u}{the cumulative gas used in the block containing the transaction receipt as of immediately after the transaction has happened, $R_\mathbf{u}$}, the set of logs created through execution of the transaction, \hyperlink{R l}{$R_\mathbf{l}$} and the Bloom filter composed from information in those logs, \hyperlink{R b}{$R_\mathbf{b}$}:
\begin{equation}
R \equiv (R_{\boldsymbol{\sigma}}, R_\mathbf{u}, \hyperlink{R b}{R_\mathbf{b}}, \hyperlink{R l}{R_\mathbf{l}})
\end{equation}

@@

We define the Bloom filter function, $M$, to reduce a log entry into a single 256-byte hash:
\begin{equation}
M(O) \equiv \hyperlink{bigvee}{\bigvee}_{t \in \{O_\mathbf{a}\} \cup O_\mathbf{t}} \big( M_{3:2048}(t) \big)
\end{equation}

where $M_{3:2048}$ is a specialised Bloom filter that sets three bits out of 2048, given an arbitrary byte sequence. It does this through taking the low-order 11 bits of each of the first three pairs of bytes in a Keccak-256 hash of the byte sequence.\footnote{11 bits $= 2^{2048}$, and the low-order 11 bits is the modulo 2048 of the operand, which is in this case is "each of the first three pairs of bytes in a Keccak-256 hash of the byte sequence".} Formally:
\begin{eqnarray}
M_{3:2048}(\mathbf{x}: \mathbf{x} \in \mathbb{B}) & \equiv & \mathbf{y}: \mathbf{y} \in \mathbb{B}_{256} \quad \text{where:}\\
\mathbf{y} & = & (0, 0, ..., 0) \quad \text{except:}\\
\forall_{i \in \{0, 2, 4\}}&:& \mathcal{B}_{m(\mathbf{x}, i)}(\mathbf{y}) = 1\\
m(\mathbf{x}, i) &\equiv& \mathtt{\tiny KEC}(\mathbf{x})[i, i + 1] \bmod 2048
\end{eqnarray}

@@

We can assert a block's validity if and only if it satisfies several conditions: it must be internally consistent with the ommer and transaction block hashes and the given transactions $B_\mathbf{T}$ (as specified in sec \ref{ch:finalisation}), when executed in order on the base state $\boldsymbol{\sigma}$ (derived from the final state of the parent block), result in a new state of the identity $H_\mathbf{r}$:
\begin{equation}
\begin{array}[t]{lclc}\hypertarget{H r}{
H_\mathbf{r} &\equiv& \mathtt{\small \hyperlink{trie}{TRIE}}(L_S(\hyperlink{Pi}{\Pi}(\boldsymbol{\sigma}, B))) & \wedge }\hypertarget{H o}{\\
H_\mathbf{o} &\equiv& \mathtt{\small KEC}(\mathtt{\small RLP}(L_H^*(\hyperlink{B U}{B_\mathbf{U}}))) & \wedge }\hypertarget{H t}{\\
H_\mathbf{t} &\equiv& \mathtt{\small TRIE}(\{\forall i < \lVert B_\mathbf{T} \rVert, i \in \mathbb{P}: p(i, L_\mathbf{T}(B_\mathbf{T}[i]))\}) }& \wedge \hypertarget{H e}{\\
H_\mathbf{e} &\equiv& \mathtt{\small TRIE}(\{\forall i < \lVert \hyperlink{Transaction Receipt}{B_\mathbf{R}} \rVert, i \in \mathbb{P}: p(i, L_R(B_\mathbf{R}[i]))\}) & \wedge }\hypertarget{H b}{\hypertarget{r}{\\
H_\mathbf{b} &\equiv& \hyperlink{bigvee}{\bigvee}_{\mathbf{r} \in B_\mathbf{R}} \big( \mathbf{r}_\mathbf{b} \big)}}
\end{array}
\end{equation}

@@

The values stemming from the computation of transactions, specifically the \hyperlink{transaction Receipt}{transaction receipts, $B_\mathbf{R}$}, and that defined through the transaction's \hyperlink{Pi}{stateR-accumulation function, $\Pi$}, are formalised later in section \ref{sec:statenoncevalidation}.

@@

The nonce of a block, $\hyperlink{H n}{H_\mathbf{n}}$, must satisfy the relations:
\begin{equation}
n \leqslant \frac{2^{256}}{\hyperlink{H d}{H_\mathbf{d}}} \quad \wedge \quad m = \hyperlink{H m}{H_\mathbf{m}}
\end{equation}
with $(n, m) = \mathtt{PoW}(H_{\hcancel{n}}, \hyperlink{H n}{H_\mathbf{n}}, \mathbf{d})$.

@@

Thus we are able to define the \hypertarget{BlockHeaderValidityFunction}{block header validity function $V(H)$}:\label{eq:BlockHeaderValidityFunction}
\begin{eqnarray}
V(H) & \equiv &  n \leqslant \frac{2^{256}}{\hyperlink{H d}{H_\mathbf{d}}} \wedge m = \hyperlink{H m}{H_\mathbf{m}} \quad \wedge \\
& & \hyperlink{H d}{H_\mathbf{d}} = D(H) \quad \wedge \\
& & \hyperlink{H g}{H_\mathbf{g}} \le \hyperlink{H l}{H_\mathbf{l}}  \quad \wedge \\
& & \hyperlink{H l}{H_\mathbf{l}} < {P(H)_H}_\mathbf{l} + \left\lfloor\frac{{P(H)_H}_\mathbf{l}}{1024}\right\rfloor  \quad \wedge \\
& & \hyperlink{H l}{H_\mathbf{l}} > {P(H)_H}_\mathbf{l} - \left\lfloor\frac{{P(H)_H}_\mathbf{l}}{1024}\right\rfloor  \quad \wedge \\
& & \hyperlink{H l}{H_\mathbf{l}} \geqslant 125000  \quad \wedge \\
& & \hyperlink{H s}{H_\mathbf{s}} > {P(H)_H}_\mathbf{s} \quad \wedge \\
& & \hyperlink{H i}{H_\mathbf{i}} = {P(H)_H}_\mathbf{i} +1 \quad \wedge \\
& & \lVert \hyperlink{H x}{H_\mathbf{x}} \rVert \le 32
\end{eqnarray}
where $(n, m) = \mathtt{PoW}(H_{\hyperlink{h cancel n}{\hcancel{n}}}, \hyperlink{H n}{H_\mathbf{n}}, \mathbf{d})$

@@

The canonical blockchain is a path from root to leaf through the entire block tree. In order to have consensus over which path it is, conceptually we identify the path that has had the most computation done upon it, or, the \textit{heaviest} path. Clearly one factor that helps determine the heaviest path is the block number of the leaf, equivalent to the number of blocks, not counting the unmined \hyperlink{GenesisBlock}{genesis block}, in the path. The longer the path, the greater the total mining effort that must have been done in order to arrive at the leaf. This is akin to existing schemes, such as that employed in Bitcoin-derived protocols.

@@

\hypertarget{Gamma}{
We may now define the function, $\Gamma$, that maps a block $B$ to its initiation state:
\begin{equation}
\Gamma(B) \equiv \begin{cases}
\boldsymbol{\sigma}_0 & \text{if} \quad P(B_H) = \varnothing \\
\boldsymbol{\sigma}_\mathbf{i}: \mathtt{\small \hyperlink{trie}{TRIE}}(L_S(\boldsymbol{\sigma}_\mathbf{i})) = {P(B_H)_H}_\mathbf{r} & \text{otherwise}
\end{cases}
\end{equation}
}
Here, $\mathtt{\small TRIE}(L_S(\boldsymbol{\sigma}_\mathbf{i}))$ means the hash of the root node of a trie of state $\boldsymbol{\sigma}_\mathbf{i}$; it is assumed that implementations will store this in the state database, which is trivial and efficient since the trie is by nature an immutable data structure.
\hypertarget{Phi}{
And finally we define $\Phi$, the block transition function, which maps an incomplete block $B$ to a complete block $B'$:}
\begin{eqnarray}
\Phi(B) & \equiv & B': \quad B' = B^* \quad \text{except:} \\
B'_\mathbf{n} & = & n: \quad x \leqslant \frac{2^{256}}{\hyperlink{H d}{H_\mathbf{d}}} \\
B'_\mathbf{m} & = & m \quad \text{with } (x, m) = \mathtt{PoW}(B^*_{\hyperlink{h cancel n}{\hcancel{n}}}, n, \mathbf{d}) \\
B^* & \equiv & B \quad \text{except:} \quad {\hyperlink{Transaction Receipt}{B^*_\mathbf{r}}} = \hyperlink{r}{r}(\hyperlink{Pi}{\Pi}(\Gamma(B), B))
\end{eqnarray}
With $\mathbf{d}$ being a dataset as specified in appendix \ref{app:ethash}.

@@

In the case of \hyperlink{Transaction Receipt}{${B_\mathbf{R}}[n]_\mathbf{u}$}, we take a similar approach defining each item as the gas used in evaluating the corresponding transaction summed with the previous item (or zero, if it is the first), giving us a running total:
\begin{equation}
\mathbf{R}[n]_\mathbf{u} = \begin{cases} 0 & \text{if} \quad n < 0 \\
\begin{array}[b]{l}
\hyperlink{Upsilong}{\Upsilon^g}(\mathbf{R}[n - 1]_{\boldsymbol{\sigma}}, B_\mathbf{T}[n])\\ \quad + \mathbf{R}[n-1]_\mathbf{u}
\end{array}
 & \text{otherwise} \end{cases}
\end{equation}

@@

\hypertarget{Pi}{
Finally, we define \hyperlink{Pi}{$\Pi$} as the new state given the block reward function \hyperlink{Omega}{$\Omega$} applied to the final transaction's resultant state, $\hyperlink{ell}{\ell}(B_\mathbf{R})_{\boldsymbol{\sigma}}$:
\begin{equation}
\Pi(\boldsymbol{\sigma}, B) \equiv \hyperlink{Omega}{\Omega}(B, \ell(\hyperlink{R sigma}{\mathbf{R})_{\boldsymbol{\sigma}}})
\end{equation}}

@@

One plague of the Bitcoin world is ASICs. These are specialised pieces of compute hardware that exist only to do a single task (\cite{ASICSmithMJS}, \cite{ASICWiki}). In Bitcoin's case the task is the SHA256 hash function (\cite{ASICbitcoinwiki}). While ASICs exist for a proof-of-work function, both goals are placed in jeopardy. Because of this, a proof-of-work function that is ASIC-resistant (i.e. difficult or economically inefficient to implement in specialised compute hardware) has been identified as the proverbial silver bullet.

Two directions exist for ASIC resistance; firstly make it sequential memory-hard, i.e. engineer the function such that the determination of the \hyperlink{block nonce}{nonce} requires a lot of memory and bandwidth such that the memory cannot be used in parallel to discover multiple nonces simultaneously. The second is to make the type of computation it would need to do general-purpose; the meaning of ``specialised hardware''  for a general-purpose task set is, naturally, general purpose hardware and as such commodity desktop computers are likely to be pretty close to ``specialised hardware'' for the task. For Ethereum \textit{Frontier} and \textit{Homestead} we have chosen the first path.

More formally, the proof-of-work function takes the form of $\mathtt{PoW}$:
\begin{equation}
m = \hyperlink{H m}{H_\mathbf{m}} \quad \wedge \quad n \leqslant \frac{2^{256}}{\hyperlink{H d}{H_\mathbf{d}}} \quad \text{with} \quad (m, n) = \mathtt{PoW}(H_{\hyperlink{h cancel n}{\hcancel{n}}}, H_\mathbf{n}, \mathbf{d})
\end{equation}

Where $H_{\hyperlink{h cancel n}{\hcancel{n}}}$ is the new block's header but \textit{without} the nonce and mix-hash components; $\hyperlink{H n}{H_\mathbf{n}}$ is the nonce of the block header; $\mathbf{d}$ is a large data set needed to compute the mixHash and $\hyperlink{H d}{H_\mathbf{d}}$ is the new block's difficulty value (i.e. the block difficulty from section \ref{ch:ghost}). $\mathtt{PoW}$ is the proof-of-work function which evaluates to an array with the first item being the mixHash and the second item being the \hyperlink{block nonce}{block nonce}, a pseudo-random number cryptographically dependent on $H$ and $\mathbf{d}$. The underlying algorithm is called Ethash and is described below.
\subsubsection{Ethash}
Ethash is the PoW algorithm for Ethereum \textit{Frontier} and \textit{Homestead}. It is the latest version of Dagger-Hashimoto, introduced by \cite{dagger} and \cite{hashimoto}, although it can no longer appropriately be called that since many of the original features of both algorithms were drastically changed with R\&D from February 2015 until May 4 2015 (\cite{commitdateforEthash}). The general route that the algorithm takes is as follows:
